---
title: "Binomial Distribution"
author: "Sreeda Vemuri"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

This assignment will introduce you to the binomial distribution.  First, type your full name in the `author` field above.

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# be sure to run lines of code like this to load code libraries before running subsequent code
library("ggplot2")
```

Hint: the keyboard shortcut to create a code-block environment is `CTRL-ALT-I` (`CMD-ALT-I` on a Mac).

---

## Commenting

Inside the code blocks, the typed information after the pound key (#) are called *comments*.  Programmers use comments to remind themselves and teammates what each section of code does.

Run the following chunk of code, try to start to understand how the numbers, variables, and syntax go together.  Write (at least) 5 comments inside the code block to demonstrate your understanding.

```{r}
# this is an example of a comment
n <- 10 # sample size
k <- 5 # median
p <- 0.32 # probability

kvals <- 0:n
tf <- kvals == k
probabilities <- dbinom(kvals, n, p)
df <- data.frame(kvals, tf, probabilities) 

custom_subtitle <- paste0("n = ", n, ", ",
                          "k = ", k, ", ",
                          "p =  ", p, "; ",
                          "P(k = ", k,
                          ") = ", round(dbinom(k,n,p), 4)) # assigns letters for the variables, and also assigns a value to a function

ggplot(df, aes(x = kvals, y = probabilities, fill = tf)) + #creates a graph to plot the data above
  geom_bar(stat = "identity") +
  labs(title = "Binomial Distribution",
       subtitle = custom_subtitle,
       caption = "Math 32",
       x = "k values",
       y = "probability") 
```

---

## Sample Size

Keep the parameters $k = 5$ and $p = 0.32$, but build a couple of graphs for a couple of different values of $n$, ($0 < k < n$), then write a sentence or two about what you observe with the graphs.

### Sample Size Graph 1
```{r}
# this is an example of a comment
n <- 7 # sample size
k <- 5 # True Value
p <- 0.32 # probability

kvals <- 0:n
tf <- kvals == k
probabilities <- dbinom(kvals, n, p)
df <- data.frame(kvals, tf, probabilities) 

custom_subtitle <- paste0("n = ", n, ", ",
                          "k = ", k, ", ",
                          "p =  ", p, "; ",
                          "P(k = ", k,
                          ") = ", round(dbinom(k,n,p), 4)) # assigns letters for the variables, and also assigns a value to a function

ggplot(df, aes(x = kvals, y = probabilities, fill = tf)) + #creates a graph to plot the data above
  geom_bar(stat = "identity") +
  labs(title = "Binomial Distribution",
       subtitle = "Sample Size - 1",
       caption = "Math 32",
       x = "k values",
       y = "probability") 
```

```{r, echo = TRUE, eval = FALSE}

This graph changes the number of bars in the graph making the "TRUE Value" slightly askew to the median.

```

This graph changes the number of bars in the graph making the "TRUE Value" slightly askew to the median.

### Sample Size Graph 2
```{r}
# this is an example of a comment
n <- 15 # sample size
k <- 5 # True Value
p <- 0.32 # probability

kvals <- 0:n
tf <- kvals == k
probabilities <- dbinom(kvals, n, p)
df <- data.frame(kvals, tf, probabilities) 

custom_subtitle <- paste0("n = ", n, ", ",
                          "k = ", k, ", ",
                          "p =  ", p, "; ",
                          "P(k = ", k,
                          ") = ", round(dbinom(k,n,p), 4)) # assigns letters for the variables, and also assigns a value to a function

ggplot(df, aes(x = kvals, y = probabilities, fill = tf)) + #creates a graph to plot the data above
  geom_bar(stat = "identity") +
  labs(title = "Binomial Distribution",
       subtitle = "Sample Size - 2",
       caption = "Math 32",
       x = "k values",
       y = "probability") 
```
```{r, echo = TRUE, eval = FALSE}

This graph also Changes the number of bar, but the k = 5 makes the true value the median of the numbers

```


---

## Items Observed

Keep the parameters $n = 10$ and $p = 0.32$, but build a couple of graphs for a couple of different values of $k$, ($0 < k < n$), then write a sentence or two about what you observe with the graphs.

### Items Observed Graph 1
```{r}
# this is an example of a comment
n <- 10 # sample size
k <- 3 # True Value
p <- 0.32 # probability

kvals <- 0:n
tf <- kvals == k
probabilities <- dbinom(kvals, n, p)
df <- data.frame(kvals, tf, probabilities) 

custom_subtitle <- paste0("n = ", n, ", ",
                          "k = ", k, ", ",
                          "p =  ", p, "; ",
                          "P(k = ", k,
                          ") = ", round(dbinom(k,n,p), 4)) # assigns letters for the variables, and also assigns a value to a function

ggplot(df, aes(x = kvals, y = probabilities, fill = tf)) + #creates a graph to plot the data above
  geom_bar(stat = "identity") +
  labs(title = "Binomial Distribution",
       subtitle = "Items Observed - 1",
       caption = "Math 32",
       x = "k values",
       y = "probability") 
```

```{r, echo = TRUE, eval = FALSE}

I now realize that K = 3 is just highlighting the bar number 3, and causing the True Value to be at 3 and there's no other significance to that value. 

```

### Items Observed Graph 2

```{r}
# this is an example of a comment
n <- 10 # sample size
k <- 8 # True Value
p <- 0.32 # probability

kvals <- 0:n
tf <- kvals == k
probabilities <- dbinom(kvals, n, p)
df <- data.frame(kvals, tf, probabilities) 

custom_subtitle <- paste0("n = ", n, ", ",
                          "k = ", k, ", ",
                          "p =  ", p, "; ",
                          "P(k = ", k,
                          ") = ", round(dbinom(k,n,p), 4)) # assigns letters for the variables, and also assigns a value to a function

ggplot(df, aes(x = kvals, y = probabilities, fill = tf)) + #creates a graph to plot the data above
  geom_bar(stat = "identity") +
  labs(title = "Binomial Distribution",
       subtitle = "Items Observed - 2",
       caption = "Math 32",
       x = "k values",
       y = "probability") 
```

```{r, echo = TRUE, eval = FALSE}

K = 8, makes it so that there a slight chance to get the true value, because it is not close to the median at all. 

```




---

## Prior Probability

Keep the parameters $n = 10$ and $k = 5$, but build a couple of graphs for a couple of different values of $p$, ($0 < p < 1$), then write a sentence or two about what you observe with the graphs.

### Prior Probability Graph 1
```{r}
# this is an example of a comment
n <- 10 # sample size
k <- 5 # median
p <- 0.5 # probability

kvals <- 0:n
tf <- kvals == k
probabilities <- dbinom(kvals, n, p)
df <- data.frame(kvals, tf, probabilities) 

custom_subtitle <- paste0("n = ", n, ", ",
                          "k = ", k, ", ",
                          "p =  ", p, "; ",
                          "P(k = ", k,
                          ") = ", round(dbinom(k,n,p), 4)) # assigns letters for the variables, and also assigns a value to a function

ggplot(df, aes(x = kvals, y = probabilities, fill = tf)) + #creates a graph to plot the data above
  geom_bar(stat = "identity") +
  labs(title = "Binomial Distribution",
       subtitle = "Prior Probability - 1",
       caption = "Math 32",
       x = "k values",
       y = "probability") 

```

```{r, echo = TRUE, eval = FALSE}

The curve appears to be symmetric, and the probability of the event occuring is split perfectly. 

```

### Prior Probability Graph 2

```{r}
# this is an example of a comment
n <- 10 # sample size
k <- 5 # median
p <- 0.8 # probability

kvals <- 0:n
tf <- kvals == k
probabilities <- dbinom(kvals, n, p)
df <- data.frame(kvals, tf, probabilities) 

custom_subtitle <- paste0("n = ", n, ", ",
                          "k = ", k, ", ",
                          "p =  ", p, "; ",
                          "P(k = ", k,
                          ") = ", round(dbinom(k,n,p), 4)) # assigns letters for the variables, and also assigns a value to a function

ggplot(df, aes(x = kvals, y = probabilities, fill = tf)) + #creates a graph to plot the data above
  geom_bar(stat = "identity") +
  labs(title = "Binomial Distribution",
       subtitle = "Prior Probability - 2",
       caption = "Math 32",
       x = "k values",
       y = "probability") 
```


```{r, echo = TRUE, eval = FALSE}

The probability of the event occurring, according to the graph is 80%, making the graph askew to the right.

```




---

# Epilogue

You should have 7 graphs in total and (at least) 3 sentences of your observations about the results.

Save your work in the "Math 32_-_Exploration 2_-_Binomial Distribution_-_your name.rmd" convention, `knit` the code, and then upload the HTML file here back into our CatCourses page.
