---
title: "Introduction to R Markdown"
author: "Sreeda Vemuri"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

This assignment will introduce you to the `R Markdown` coding language.  Moreover, this `R Markdown` file (note the *.rmd extension) is a setting called "literate programming" where a user can alternate between regular typing---enhanced with markdown---and computer code (`R` or `Python`).  For reference, here is RStudio's own [cheat sheet for R Markdown](https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)

---

Type your full name in the `author` field above.  State your major (including emphasis if applicable) and list your classes---course number and course name---below as an ***ordered-list***, from the one you think will be the most difficult to the least difficult.

##### Major: Computer Science Engineering with Applied Mathematics Minor

##### Classes:
1. Math 032 - Statistics and Probability
2. Bio 005 - Biology Today
3. GASP 010 - Painting



---

Why are you enrolled in Math 32? What are you hoping to learn and gain? (Please supply an answer that is more in-depth than "This course fulfills my statistics requirement".)

In Math 032, I plan on learning about probability, and maybe it will help me in the long run with stocks and maybe even gambling once in a while. And also trying to make sense of how to understand the different statistics and learn about the validity of the data spreads provided by the news. 

---

In 2-3 sentences, type a science fact that you find fascinating into the code block.  Set the parameters `echo = TRUE` and `eval = FALSE` so that RStudio does not treat the following as R code.

```{r, echo = TRUE, eval = FALSE}
Stomach Acid can burn trhough steel because of the acidity of Hydrocloric Acid. The PH level is about 2.3 of HydroCloric Acid. And for this reason, if a person does not eat for longs periods of time, in severe causes it causes ulcers in th stomach and throat and in non severe cases, it just causes acidity. 
```

---

Spend about 5 to 10 minutes typing out your answer to "What are you hopes and goals?" as a ***bullet-point list***.  Do not worry about grammar; just type fast.

- Good Grades

- Understanding Material

- Balancing School and sleep

* (Trying to be social in the Pandemic)

* Helping Mom out

* Interships
* Resume
* GMAT prep
* Passport renewal
* Apartment hunt -FAll
* Email professors- remote learning??
* Laundry
* Future-4bed House
* Also future- small backyard
* MBA in New York
* Car-aventador S
* Build PC before Fall 2020
* JOB HUNT
* Scholarships?
* internship remote possibilites? 
* Be respectable at job
* Cool closet
* i honestly don't know anymore


---

Spend about 5 minutes typing out your answer to "What are your fears and obstacles?" into the code block.  Set the parameters `echo = FALSE` so that you do not show this part to your teachers and `eval = FALSE` so that RStudio does not treat the following as R code.

```{r, echo = FALSE, eval = FALSE}
*Opening up
*crying in public
*Being abandoned
*Being homeless
*beliveing someone and getting my heart broken. 

```

---

For this semester, instructor Derek wants to assemble a class playlist of popular music.  What is a good, lively song to add to a class playlist?  Place the song title between asterisks to *italicize* the song title and place the artist between two-pairs of asterisks so that the artist name appears in **boldface**.  (Beyond profanity, please avoid offensive content.)

*I Like Me Better* -- **Lauv**


---

Save your work in the "Introduction_-_your name.rmd" convention, `knit` the code, and then upload the HTML file here back into our CatCourses page.
