---
title: "Math 32:  Exploration 4"
author: "Sreeda Vemuri"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
        theme: cerulean
---

In this `exploration`, you are asked to

* generate random numbers from continuous distributions
* observe the visualizations of those number sets

1. First,
    a. Type your name into the author field in the YAML parameters above
    b. Also save this file while replacing "template" with your full name (be sure to keep the *.rmd file format).

```{r, message = FALSE, warning = FALSE}
library("ggplot2")
```

# Uniform Distribution

2. Demonstrate your understanding about the example code block below by adding at least 5 comments in the code block in your own words.

```{r}
N <- 1000
a <- 1
b <- 32
data <- runif(N, a, b) #function that is used when creating the plots
df <- data.frame(data)

ggplot(df, aes(y = data)) +
  coord_flip() +
  geom_boxplot() + #Creates a boxplot
  labs(title = paste0("Uniform Distribution U(", a, ",", b, ")"),
       subtitle = "boxplot",
       caption = "Sreeda Vemuri") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  ylim(0,50) 

ggplot(df, aes(x = data)) +
  geom_density(kernel = "gaussian") + # curve that plots the function in a linear format
  labs(title = paste0("Uniform Distribution U(", a, ",", b, ")"),
       subtitle = "density plot",
       caption = "Sreeda Vemuri") +
  theme(panel.background = element_blank(), #color of the background of the bar plot, can be changed by element_rect(fill, color, size)
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  xlim(0,50) #x-axis limits (0-50)
```

3. Repeat step 2 in a new code block, but with different values of $a$ and $b$ in $0 < a < b < 50$

```{r}
N <- 1000
a <- 20
b <- 43
data <- runif(N, a, b)
df <- data.frame(data)

ggplot(df, aes(y = data)) +
  coord_flip() +
  geom_boxplot() +
  labs(title = paste0("Uniform Distribution U(", a, ",", b, ")"),
       subtitle = "boxplot",
       caption = "Sreeda Vemuri") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  ylim(0,50)

ggplot(df, aes(x = data)) +
  geom_density(kernel = "gaussian") +
  labs(title = paste0("Uniform Distribution U(", a, ",", b, ")"),
       subtitle = "density plot",
       caption = "Sreeda Vemuri") +
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  xlim(0,50)
```

4. Make at least two observations about the results (the graphs) in an unordered list.

* The Mean of the Boxplots compared has changed a lot. 
* the gaussian curve seems to be more compact relative to the orginal uniform distribution

# Normal Distribution

5. Make adjustments to the code below to repeat what was done above with the uniform distribution, but this time looking at the normal distribution (hint: use the `rnorm(N, mu, sd)` command).

```{r}
N <- 1000
mu <- 25
sd <- 3
data <- rnorm(N, mu, sd)
df <- data.frame(data)

ggplot(df, aes(y = data)) +
  coord_flip() +
  geom_boxplot() +
  labs(title = paste0("Normal Distribution N(", mu, ",", sd*sd, ")"),
       subtitle = "boxplot",
       caption = "Sreeda Vemuri") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  ylim(0,50)

ggplot(df, aes(x = data)) +
  geom_density(kernel = "gaussian") +
  labs(title = paste0("Normal Distribution N(", mu, ",", sd*sd, ")"),
       subtitle = "density plot",
       caption = "Sreeda Vemuri") +
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  xlim(0,50)
```

6. Repeat step 5 in a new code block, but with different values of $\mu$ and $\sigma$ in $0 < \sigma < \mu < 50$
```{r}
N <- 1000
mu <- 39
sd <- 17
data <- rnorm(N, mu, sd)
df <- data.frame(data)

ggplot(df, aes(y = data)) +
  coord_flip() +
  geom_boxplot() +
  labs(title = paste0("Normal Distribution N(", mu, ",", sd*sd, ")"),
       subtitle = "boxplot",
       caption = "Sreeda Vemuri") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  ylim(0,50)

ggplot(df, aes(x = data)) +
  geom_density(kernel = "gaussian") +
  labs(title = paste0("Normal Distribution N(", mu, ",", sd*sd, ")"),
       subtitle = "density plot",
       caption = "Sreeda Vemuri") +
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  xlim(0,50)
```
7. Make at least two observations about the results (the graphs) in an unordered list.
* The standard deviation shown in the gaussian curve in the second distribution is far greater than the first
* The boxplot for the first distribution shows the outliers unlike the second

# Exponential Distribution

8. Make adjustments to the code below to repeat what was done above with the uniform distribution, but this time looking at the exponential distribution with $\lambda = \frac{1}{\mu}$ (hint: use the `rexp(N, rate)` command).

```{r}
N <- 1000
mu <- 32
data <- rexp(N, 1/mu)
df <- data.frame(data)

ggplot(df, aes(y = data)) +
  coord_flip() +
  geom_boxplot() +
  labs(title = paste0("Exponential Distribution E(1/", mu, ")"),
       subtitle = "boxplot",
       caption = "Sreeda Vemuri") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  ylim(0,50)

ggplot(df, aes(x = data)) +
  geom_density(kernel = "gaussian") +
  labs(title = paste0("Exponential Distribution E(1/", mu, ")"),
       subtitle = "density plot",
       caption = "Sreeda Vemuri") +
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  xlim(0,50)
```

9. Repeat step 8 in a new code block, but with different a value of $\mu$ in $0 < \mu < 50$

```{r}
N <- 1000
mu <- 19
data <- rexp(N, 1/mu)
df <- data.frame(data)

ggplot(df, aes(y = data)) +
  coord_flip() +
  geom_boxplot() +
  labs(title = paste0("Exponential Distribution E(1/", mu, ")"),
       subtitle = "boxplot",
       caption = "Sreeda Vemuri") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  ylim(0,50)

ggplot(df, aes(x = data)) +
  geom_density(kernel = "gaussian") +
  labs(title = paste0("Exponential Distribution E(1/", mu, ")"),
       subtitle = "density plot",
       caption = "Sreeda Vemuri") +
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  xlim(0,50)
```

10. Make at least two observations about the results (the graphs) in an unordered list.

* the exponential distribution plots are literally just that, the gaussian curve shows an exponential decay in the distribution

* the exponential decay is also evident in the boxplot, but it is just showing that the numbers are more clustered closer to the given $\mu$

11.  Verify that you have 12 graphs in total (6 boxplots and 6 density plots), and replace "Math 32" with your full name in all of the graphs.

12. Finally, `knit` your work as either an HTML or PDF document, and upload that document back into the CatCourses assignment.