---
title: "Exploration 6:  Estimators"
author: "Sreeda Vemuri"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r, message = FALSE, warning = FALSE}
library("ggplot2")

# This randomly generates 25 random numbers from an U(0,1) distribution
one_sample <- runif(25, 0, 1)

# Note that the population variance formula does not have "n - 1"
bad_var <- function(x){
  N <- length(x)
  sigma_squared <- sum( (x - mean(x))^2 ) / N
  sigma_squared #return this number
}

# Note that the sample variance formula does have "n - 1"
good_var <- function(x){
  n <- length(x)
  s_squared <- sum( (x - mean(x))^2 ) / (n-1)
  s_squared #return this number 
}
```

## Example

Here we will run a simulation to empirically show that the *population mean*
$$\mu = \displaystyle\frac{1}{N} \sum_{i = 1}^{N} x_{i}$$
is **unbiased** when applied to samples over the $U(0,1)$ uniform distribution (whose true population mean is $\mu = \displaystyle\frac{1}{2}$).

```{r, eval = TRUE}
true_mean <- 1/2

simulation <- replicate(10000, mean( runif(25, 0, 1) ))
df <- data.frame(simulation)

ggplot(df, aes(x = simulation)) +
  geom_density(color = "blue") + 
  geom_vline(xintercept = true_mean, color = "red") +
  labs(title = "Sample Distribution for the Sample Mean",
       subtitle = "Blue: sample distribution, Red: statistic of interest",
       caption = "Math 32", 
       x = "x",
       y = "density")
```

Since the sample distribution empirically does seem to line up with the population mean, we say that the sample mean $\bar{x}$ is *unbiased*.

---

Repeat the simulation above, but with the an $U(0,32)$ distribution and the following estimators:

1.  population variance `bad_var()`

```{r, eval = TRUE}
true_mean <- 1/2

simulation <- replicate(10000, bad_var( runif(25,0,32) ))
df <- data.frame(simulation)

ggplot(df, aes(x = simulation)) +
  geom_density(color = "blue") + 
  geom_vline(xintercept = true_mean, color = "red") +
  labs(title = "Sample Distribution for the Sample Mean",
       subtitle = "Blue: sample distribution, Red: statistic of interest",
       caption = "Sreeda Vemuri", 
       x = "x",
       y = "density")
```

2.  sample variance `good_var()`
```{r, eval = TRUE}
true_mean <- 1/2

simulation <- replicate(10000, good_var(runif(25,0,32) ))
df <- data.frame(simulation)

ggplot(df, aes(x = simulation)) +
  geom_density(color = "blue") + 
  geom_vline(xintercept = true_mean, color = "red") +
  labs(title = "Sample Distribution for the Sample Mean",
       subtitle = "Blue: sample distribution, Red: statistic of interest",
       caption = "Sreeda Vemuri", 
       x = "x",
       y = "density")
```

3.  the sample standard deviation from `good_var()`
```{r, eval = TRUE}
true_mean <- 1/2

simulation <- replicate(10000, sqrt(good_var(runif(25,0,32) )))
df <- data.frame(simulation)

ggplot(df, aes(x = simulation)) +
  geom_density(color = "blue") + 
  geom_vline(xintercept = true_mean, color = "red") +
  labs(title = "Sample Distribution for the Sample Mean",
       subtitle = "Blue: sample distribution, Red: statistic of interest",
       caption = "Sreeda Vemuri", 
       x = "x",
       y = "density")
```



After each simulation, make an educated guess as to whether each estimator is biased or unbiased.

---

Verify that you have 5 graphs in total and replace "Math 32" with your full name in all of the graphs.

Finally, `knit` your work as either an HTML or PDF document, and upload that document back into the CatCourses assignment.
